---
title: "R Bridge Course Week 2 Assignment"
author: "Ann Liu-Ferrara"
date: "January 13, 2017"
output: html_document

---
1. Use the summary function to gain an overview of the data set. Then display the mean and median for at least two attributes.

```{r}
library(RCurl)
library(ggplot2)
library(plyr)
library(reshape2)
library(extrafont)

weightLoss.data <- getURL("https://raw.githubusercontent.com/ann2014/CUNY/master/WeightLoss.csv")
weightLoss.data <- read.csv(text = weightLoss.data)
head(weightLoss.data)
summary(weightLoss.data)

```

2. Create a new data frame with a subset of the columns and rows. Make sure to rename it.

```{r, echo=TRUE}
str(weightLoss.data)
new.df <- subset(weightLoss.data, select = -X)
head(new.df)
```

3. Create new column names for the new data frame.

```{r}
names(new.df)[2:4] <- c("WeightLoss_month1", "WeightLoss_month2", "WeightLoss_month3")
names(new.df)[5:7] <- c("SelfEsteem_month1", "SelfEsteem_month2", "SelfEsteem_month3")
head(new.df)
```

4. Use the summary function to create an overview of your new data frame. The print the mean and median for the same two attributes. Please compare.

```{r}
summary(new.df)
summary(weightLoss.data)
```

5. For at least 3 values in a column please rename so that every value in that column is renamed. For example, suppose I have 20 values of the letter "e" in one column. Rename those values so that all 20 would show as "excellent".

```{r}
head(new.df[, 1])
new.df[, 1] <- gsub('Control', 'Ctrl', new.df[, 1])
head(new.df[, 1])

```

6. Display enough rows to see examples of all of steps 1-5 above.

```{r}
head(new.df, 15)
```
7. BONUS - place the original .csv in a github file and have R read from the link. This will be a very useful skill as you progress in your data science education and career.

https://raw.githubusercontent.com/ann2014/CUNY/master/WeightLoss.csv
